---
title: "Stronger Together? Linked Fate and Voter Preferences in the 2020 Election"
author: "A. Jordan Nafa, Meredith Walsh Niezgoda, P. DeAnne Roark, and Valerie Martinez-Ebers"
date: "September 15th, 2022"
institute: "University of North Texas"
format: 
  revealjs:
    theme: blood
    transition: fade
    highlight-style: monokai
    toc-depth: 2
    self-contained: true
    citations-hover: true
    code-link: true
    code-block-bg: "#272822"
    css: "style.css"
    include-in-header: "math-colors.js"
    pdf-separate-fragments: true
editor: visual
width: 1360
height: 800
bibliography: "../../assets/references.bib"
csl: "../../assets/apsa.csl"
link-citations: yes
---

```{r, include=FALSE}
# Load the required libraries
pacman::p_load(
  "tidyverse",
  "arrow",
  "ggdist",
  install = FALSE
)

## Base theme for the figures
fig_theme <- function(.base_size, ...) {
  .plot_theme <- theme_light(base_size = .base_size, base_family = "serif") +
  ## Settings sepcific to the reveal.js theme
  theme(
    strip.background = element_blank(),
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background = element_rect(fill = "transparent", colour = NA),
    legend.background = element_rect(fill = "transparent", colour = NA),
    legend.key = element_rect(fill = "transparent", colour = NA),
    plot.title = element_text(color = "white", face = "bold"),
    plot.subtitle = element_text(color = "white"),
    plot.caption = element_text(color = "white", face = "italic"),
    axis.text = element_text(color = "white"),
    axis.title = element_text(color = "white", face = "bold"),
    legend.title = element_text(color = "white", face = "bold"),
    legend.text = element_text(color = "white", face = "bold"),
    strip.text = element_text(face = "bold"),
    plot.caption.position = "plot",
    legend.position = "top"
  )
  
  return(.plot_theme)
}
```

## Introduction

::: incremental
-   How do intersecting identities influence candidate preference among members of pan-ethnic racial groups?

    -   Do these relationships vary within pan-ethnic identity groups?

-   We examine differences within and between pan-ethnic identity groups

    -   Demonstrate the possibility of moving beyond "Latinos and Asian Americans are not monolithic" and treating heterogeneity among pan-ethnic groups as substantively interesting
:::

## Previous Research

::: incremental
-   Black Americans' Linked Fate [@Dawson1995]

-   Latinos' and Asian Americans' Complicated Linked Fate [@Junn2008; @Sanchez2019]

-   Intersectional Linked Fate [@Junn2017; @Gay2016]

-   Intra-group differences in Linked Fate [@Arora2021]

-   We attempt to bridge several lines of research on pan-ethnic linked fate that have thus far developed largely independent of one another
:::

## Social identities in the Era of Trump

::: incremental
-   Race, ethnicity, gender, and social identity theory

-   What makes some identities more salient than others?

-   The Trump presidency and the salience of group identity

-   Theoretical Expectations

    -   The salience of both pan-ethnic and gender-based dimensions of linked fate will vary across different sub-groups

    -   The threat against multiple identities will activate intersectional linked fate

    -   Relationship between intersectional linked fate and candidate preference is greater than the gender or pan-ethnic dimensions alone
:::

# Methods and Data

## Data

Asian (N = 3,898) and Latino (N = 3,950) Respondents on the 2020 Collaborative Multi-Racial Post-Election Survey [@Frasure2021]

::: incremental
-   Dependent Variable

    $y_{i} = \begin{cases}1 \text{ if respondent } i \text{ Supported Trump in 2020}\\ 0 \text{ Otherwise}\end{cases}$

-   Dimensions of Linked Fate: Pan-Ethnic, Gender, and Intersectional

    -   *"How much do you think what happens to the following groups here in the U.S. will have something to do with what happens in your life? What happens to \[identity group\] will have..."*

-   Sex, Citizenship, Age, Education, Ancestry, Partisan Identification, percentage University educated, Median Age
:::

## Multilevel Regression and Post-Stratification (MrP)

::: incremental
-   Bayesian Multilevel Logistic Regression

-   Partial pooling of categorical predictors to penalize model complexity and prevent over-fitting [@Gelman2007; @Park2004]

-   Weakly informative priors on the coefficients and hierarchical variances to stabilize computation [@Gelman2007; @Gelman2021]

-   For non-census variables we first develop models to predict their distribution in the population and add them to the stratification table
:::

## MrP Model Specification {.smaller}

$$
\begin{aligned}
y &\sim \mathrm{Binomial}(n, k, \theta)\\
\mathrm{logit}(\theta) &= \alpha + \beta_{1}\text{Female}_{i} + \beta_{2}\text{Citizen}_{i} + \upsilon_{j[i]}^{\mathrm{Ancestry}} + \upsilon_{l[i]}^{\mathrm{Education}} + \upsilon_{m[i]}^{\mathrm{Age}} +\\
&\quad \upsilon_{o[i]}^{\mathrm{Partisan}} + \upsilon_{p[i]}^{\mathrm{Linked~Fate}} + \upsilon_{q[i],p[i]}^{\mathrm{Female \times Linked~Fate}} + \upsilon_{q[i],j[i]}^{\mathrm{Female \times Ancestry}} +\\
&\quad \upsilon_{p[i],j[i]}^{\mathrm{Linked~Fate \times Ancestry}}\\
\text{where}\\
\upsilon_{j}^{\mathrm{Ancestry}} &\sim \mathcal{N}(\left[\gamma_{0} + \gamma_{1}\text{% College}_{j} + \gamma_{2}\text{Median Age}_{j} \right], ~\sigma^{\mathrm{Ancestry}}) \quad \text{for } j \in \{1, 2, \dots, J\}\\
\upsilon_{l}^{\mathrm{Education}} &\sim \mathcal{N}(0, ~\sigma^{\mathrm{Education}})  \quad \text{for } l \in \{1, 2, \dots, L\}\\
\upsilon_{m}^{\mathrm{Age}}&\sim  \mathcal{N}(0, ~\sigma^{\mathrm{Age}})  \quad \text{for } m \in \{1, 2, \dots, M\}\\
\upsilon_{o}^{\mathrm{Partisan}} &\sim \mathcal{N}(0, ~\sigma^{\mathrm{Partisan}})  \quad \text{for } o \in \{1, 2, \dots, O\}\\
\upsilon_{p}^{\mathrm{Linked~Fate}} &\sim \mathcal{N}(0, ~\sigma^{\mathrm{Linked~Fate}})\quad \text{for } p \in \{1, 2, \dots, P\}\\
\upsilon_{q,p}^{\mathrm{Female \times Linked~Fate}} &\sim \mathcal{N}(0, ~\sigma^{\mathrm{Female \times Linked~Fate}}) \quad \text{for } q \in \{1, 2, \dots, Q\} \text{ and } p \in \{1, 2, \dots, P\}\\
\upsilon_{q,j}^{\mathrm{Female \times Ancestry}} &\sim \mathcal{N}(0, ~\sigma^{\mathrm{Female \times Ancestry}}) \quad \text{for } q \in \{1, 2\} \text{ and } j \in \{1, 2, \dots, J\}\\
\upsilon_{p,j}^{\mathrm{Linked~Fate \times Ancestry}}&\sim \mathcal{N}(0, ~\sigma^{\mathrm{Linked~Fate \times Ancestry}}) \quad \text{for } p \in \{1, 2, \dots, P\} \text{ and } j \in \{1, 2, \dots, J\}
\end{aligned}
$$

## Post-Stratification Stage

Post-Stratified estimate by Linked Fate and Ancestry or Sex based on the 2019 American Community Survey 5-year Integrated Public Use Microdata Series data [@Ruggles2022]

$$\begin{aligned}\theta^{\mathrm{MrP}} = \frac{\sum_{p,j \in P,J}N_{p,j}\cdot \theta_{p,j}}{\sum_{p,j \in P,J}N_{p,j}}\end{aligned}$$

::: incremental
-   $\theta$ is the predicted probability of support for Trump from our model

-   $N$ is the observed cell count in the post-stratification table

-   Summing over the population predictions within each pair of cells $p,j \in P,J$ and dividing by the population total yields $\theta^{\mathrm{MrP}}$ which represents the census-adjusted parameter estimate
:::

# Results

------------------------------------------------------------------------

```{r results-by-group, echo=FALSE, dpi=300, dev.args = list(bg = 'transparent'), fig.height=10, fig.width=18, fig.align='center', cache=TRUE, dev='svg'}
combined_totals <- read_parquet(".data/combined_totals.gz.parquet")

combined_totals %>% 
  ggplot(aes(x = linked_fate_int, fill = type, y = theta)) +
  facet_wrap(~ group, ncol = 2, scales = "free_y") +
  stat_gradientinterval(
    aes(slab_alpha = stat(pdf), shape = type),
    point_interval = "median_qi",
    position = "dodge",
    fill_type = "gradient",
    .width = c(.89, 0.95),
    point_size = 3
  ) +
  # Set the shape parameter for each group
  scale_shape_manual(values = c(22, 24, 23)) +
  # Set the parameters of the y axis
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5), limits = c(0, 0.50)) +
  # Use the Viridis color pallet
  scale_fill_manual(values = c("#9400D3", "#7FFF00", "#CD0000")) +
  # Add labels to the plot
  labs(
    x = "Linked Fate (Low - High)",
    y = "Probability of Support for Trump",
    title = str_wrap("Figure 1. Post-Stratified Probabilities of Support for Trump by Linked Fate Dimension", width = 80)
  ) +
  # Setting the parameters for the plot legend
  guides(fill = guide_legend(
    title = "Dimension",
    override.aes = list(
      shape = c(22, 24, 23),
      fill = c("#9400D3", "#7FFF00", "#CD0000"),
      size = 10
    )
  ),
  slab_fill = "none",
  shape = "none",
  slab_alpha = "none"
  ) +
    fig_theme(.base_size = 30) +
  # Additional theme adjustments for title and caption alignment
  theme(
    plot.caption.position = "plot",
    plot.title.position = "plot",
    legend.position = "top",
    plot.margin = margin(2,8,5,5, "mm")
  )
```

------------------------------------------------------------------------

```{r results-by-gender-latino, echo=FALSE, dpi=300, dev.args = list(bg = 'transparent'), fig.height=10, fig.width=18, fig.align='center', cache=TRUE, dev='svg'}
combined_gender_gap <- read_parquet(".data/combined_gender_gap.gz.parquet")

combined_gender_gap  %>% 
  filter(group == "Hispanic or Latino") %>% 
  ggplot(aes(x = linked_fate_int, fill = type, y = theta)) +
  facet_wrap(~ str_c(group, ": ", female), ncol = 2, scales = "free_y") +
  stat_gradientinterval(
    aes(slab_alpha = stat(pdf), shape = type),
    point_interval = "median_qi",
    position = "dodge",
    fill_type = "gradient",
    .width = c(.89, 0.95),
    point_size = 3
  ) +
  # Set the shape parameter for each group
  scale_shape_manual(values = c(22, 24, 23)) +
  # Set the parameters of the y axis
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5), limits = c(0, 0.50)) +
  # Use the Viridis color pallet
  scale_fill_manual(values = c("#9400D3", "#7FFF00", "#CD0000")) +
  # Add labels to the plot
  labs(
    x = "Linked Fate (Low - High)",
    y = "Probability of Support for Trump",
    title = str_wrap("Figure 2A. Post-Stratified Probabilities of Support for Trump among Latinos by Linked Fate Dimension and Gender", width = 80)
  ) +
  # Setting the parameters for the plot legend
  guides(fill = guide_legend(
    title = "Dimension",
    override.aes = list(
      shape = c(22, 24, 23),
      fill = c("#9400D3", "#7FFF00", "#CD0000"),
      size = 10
    )
  ),
  slab_fill = "none",
  shape = "none",
  slab_alpha = "none"
  ) +
    fig_theme(.base_size = 30) +
  # Additional theme adjustments for title and caption alignment
  theme(
    plot.caption.position = "plot",
    plot.title.position = "plot",
    legend.position = "top",
    plot.margin = margin(2,8,5,5, "mm")
  )
```

------------------------------------------------------------------------

```{r results-by-gender-asian, echo=FALSE, dpi=300, dev.args = list(bg = 'transparent'), fig.height=10, fig.width=18, fig.align='center', cache=TRUE, dev='svg'}
combined_gender_gap  %>% 
  filter(!group == "Hispanic or Latino") %>% 
  ggplot(aes(x = linked_fate_int, fill = type, y = theta)) +
  facet_wrap(~ str_c(group, ": ", female), ncol = 2, scales = "free_y") +
  stat_gradientinterval(
    aes(slab_alpha = stat(pdf), shape = type),
    point_interval = "median_qi",
    position = "dodge",
    fill_type = "gradient",
    .width = c(.89, 0.95),
    point_size = 3
  ) +
  # Set the shape parameter for each group
  scale_shape_manual(values = c(22, 24, 23)) +
  # Set the parameters of the y axis
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5), limits = c(0, 0.50)) +
  # Use the Viridis color pallet
  scale_fill_manual(values = c("#9400D3", "#7FFF00", "#CD0000")) +
  # Add labels to the plot
  labs(
    x = "Linked Fate (Low - High)",
    y = "Probability of Support for Trump",
    title = str_wrap("Figure 2B. Post-Stratified Probabilities of Support for Trump among Asians by Linked Fate Dimension and Gender", width = 80)
  ) +
  # Setting the parameters for the plot legend
  guides(fill = guide_legend(
    title = "Dimension",
    override.aes = list(
      shape = c(22, 24, 23),
      fill = c("#9400D3", "#7FFF00", "#CD0000"),
      size = 10
    )
  ),
  slab_fill = "none",
  shape = "none",
  slab_alpha = "none"
  ) +
    fig_theme(.base_size = 30) +
  # Additional theme adjustments for title and caption alignment
  theme(
    plot.caption.position = "plot",
    plot.title.position = "plot",
    legend.position = "top",
    plot.margin = margin(2,8,5,5, "mm")
  )
```

------------------------------------------------------------------------

```{r results-by-ancestry-latino, echo=FALSE, dpi=300, dev.args = list(bg = 'transparent'), fig.height=16, fig.width=28, fig.align='center', cache=TRUE, dev='svg'}
combined_ancestry <- read_parquet(".data/combined_ancestry_delta.gz.parquet")

combined_ancestry %>% 
    filter(group == "Hispanic or Latino") %>% 
  ggplot(aes(x = theta_delta, y = fct_rev(ancestry), fill = fct_rev(ancestry))) +
  facet_wrap(~str_c(type, " Linked Fate"), ncol = 3) +
  stat_gradientinterval(
    aes(slab_alpha = stat(pdf), shape = type),
    point_interval = "mean_qi",
    position = "dodge",
    fill_type = "gradient",
    .width = c(.80, 0.95),
    point_size = 3
  ) +
  # Set the shape parameter for each group
  scale_shape_manual(values = c(22, 24, 23)) +
  # Use the Viridis color pallet
  scale_fill_manual(values = viridisLite::turbo(n = 25)) +
  # Scales for the y axis
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), limits = c(-0.13, 0.13)) +
  # Add labels to the plot
  labs(
    y = "",
    x = latex2exp::TeX(r'($\Delta Pr(\theta \, |\, Linked\,Fate,\, Ancestry)$)', bold = TRUE),
    title = str_wrap("Figure 3A. Average Post-Stratified Contrasts for the Probability of Support for Trump among Hispanics and Latinos by Ancestry and Linked Fate Dimension", width = 100)
  ) +
  # Setting the parameters for the plot legend
  guides(fill = "none",
         slab_fill = "none",
         shape = "none",
         slab_alpha = "none"
  ) +
    fig_theme(.base_size = 40) +
  # Additional theme adjustments for title and caption alignment
  theme(
    plot.caption.position = "plot",
    plot.title.position = "plot",
    legend.position = "top",
    plot.margin = margin(2,8,5,5, "mm")
  )
```

------------------------------------------------------------------------

```{r results-by-ancestry-asian, echo=FALSE, dpi=300, dev.args = list(bg = 'transparent'), fig.height=16, fig.width=28, fig.align='center', cache=TRUE, dev='svg'}
combined_ancestry <- read_parquet(".data/combined_ancestry_delta.gz.parquet")

combined_ancestry %>% 
    filter(!group == "Hispanic or Latino") %>% 
  ggplot(aes(x = theta_delta, y = fct_rev(ancestry), fill = fct_rev(ancestry))) +
  facet_wrap(~str_c(type, " Linked Fate"), ncol = 3) +
  stat_gradientinterval(
    aes(slab_alpha = stat(pdf), shape = type),
    point_interval = "mean_qi",
    position = "dodge",
    fill_type = "gradient",
    .width = c(.80, 0.95),
    point_size = 3
  ) +
  # Set the shape parameter for each group
  scale_shape_manual(values = c(22, 24, 23)) +
  # Use the Viridis color pallet
  scale_fill_manual(values = viridisLite::turbo(n = 25)) +
  # Scales for the y axis
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), limits = c(-0.13, 0.13)) +
  # Add labels to the plot
  labs(
    y = "",
    x = latex2exp::TeX(r'($\Delta Pr(\theta \, |\, Linked\,Fate,\, Ancestry)$)', bold = TRUE),
    title = str_wrap("Figure 3B. Average Post-Stratified Contrasts for the Probability of Support for Trump among Asians by Ancestry and Linked Fate Dimension", width = 90)
  ) +
  # Setting the parameters for the plot legend
  guides(fill = "none",
         slab_fill = "none",
         shape = "none",
         slab_alpha = "none"
  ) +
    fig_theme(.base_size = 40) +
  # Additional theme adjustments for title and caption alignment
  theme(
    plot.caption.position = "plot",
    plot.title.position = "plot",
    legend.position = "top",
    plot.margin = margin(2,8,5,5, "mm")
  )
```

# Conclusions

## Main Findings

::: incremental
-   People who feel more connected to other "\[men/women\] of color" were overall less likely to support Trump in 2020

-   Relationship among Asian Americans is very non-linear across each dimension of Linked Fate, and particularly among those of East Asian descent

-   Pan-Ethnic identity groups are far from monolithic and this heterogeneity is substantively interesting

-   Not much evidence of within group-gender gaps in the joint distribution of linked fate and support for Trump across all dimensions
:::

## Conclusions and Limitations

::: incremental
-   We illustrate the application of MrP, the "gold standard for subnational public opinion estimation" [@Lax2009a], need not be limited to geography

-   Broad applicability in the racial and ethnic politics literature, particularly in descriptive work but can also be used to improve estimates in experimental research [see @Gao2022]

-   Our data is purely cross-sectional and constitutes a single snapshot in time---our aims here are purely descriptive

-   We cannot speak to the question of *why* we observe these within and between group differences in the relationships between linked fate and support for Trump as that is a causal question with a causal answer
:::

## References
